---
title: "The Natural History of Hypertension in Older Adults: A Study of Two Finnish Generational Cohorts Born 20 Years Apart - systolic BP sensitivity analyses"
output:
  html_document: default
#  pdf_document: default
classoption: landscape
date: "`r Sys.Date()`"
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Load libraries
library(pacman)

# Load the rest
p_load(psych, dplyr, knitr, kableExtra, here, lubridate, tidyverse, janitor, writexl)

# Format the kable styling
bs_style <- c("striped", "bordered")

options(kable_styling_bootstrap_options = bs_style)

```
<br><br><br><br>



# Introduce a special function that stylizes p-values

The following code chunk introduces a special function that stylizes p-values.


<br><br><br>
```{r stylize_p}
 stylize_p <- function(x) {
  case_when(
    x < 0.001 ~ "<.001",
    x >= 0.001 & x < 0.005 ~ as.character(format(round_half_up(x, 3), nsmall = 3)),
    
    x >= 0.005 & x < 0.045 ~ as.character(format(round_half_up(x, 2), nsmall = 2)),
    x >= 0.045 & x < 0.055 ~ as.character(format(round_half_up(x, 3), nsmall = 3)),
    x >= 0.055 ~ as.character(format(round_half_up(x, 2), nsmall = 2)),
    TRUE ~ "other"
  )
}



```
<br><br><br>
  



# Introduce a special function that presents mean ± sd

The following code chunk introduces a special function for mean ± sd.


<br><br><br>
```{r meansd}

 # Function to calculate mean ± sd
mean_sd <- function(data, na.rm = TRUE) {
  the_mean <- mean(data, na.rm = na.rm)
  the_sd <- sd(data, na.rm = na.rm)
  combined <- paste0(round_half_up(the_mean, digits = 1), "±", round_half_up(the_sd, digits = 1))
  return(combined)
}


```
<br><br><br>
  





# Load the datasets with the {here} package
<br><br>



```{r here_package}

# Tell the "here" package where you are right now - i.e. tell the location of the file you are running right now

here::i_am("article1/README.R")

# After that, you can check where the project root is by running "here()"

here()


#  Load the datasets using {here} package
article1tuva_file <- here("data", "a1_sensitivity_tuva.rds")
article1utuva_file <- here("data", "a1_sensitivity_utuva.rds")

article1tuva <- readRDS(article1tuva_file)
article1utuva <- readRDS(article1utuva_file)

# Clean up variables
rm(article1tuva_file, article1utuva_file)

```





# Exclusion of participants in the baseline without data for SBP and/or DBP
<br><br>

We will exclude those who don't have both an SBP and a DBP reading in the baseline.

TUVA1:


```{r exclusion_tuva1, echo=FALSE}

# TUVA1
article1tuva <- article1tuva %>% drop_na(tuva1sbp, tuva1dbp)
article1tuva %>% nrow()


```


UTUVA1:

```{r exclusion_utuva1, echo=FALSE}
# UTUVA1
article1utuva <- article1utuva %>% drop_na(utuva1sbp, utuva1dbp)
article1utuva %>% nrow()

```



  




# Count the n numbers of available BP readings for Table 2
<br><br>

Store these into a vector and display the values.

<br><br><br>
```{r n_numbers}


# TUVA1

tuva1_n <- sum(!is.na(article1tuva$tuva1sbp))
tuva1_n

psych::describe(article1tuva$tuva1sbp)

# TUVA2
tuva2_n <- sum(!is.na(article1tuva$tuva2sbp))
tuva2_n

psych::describe(article1tuva$tuva2sbp)

#TUVA3
tuva3_n <- sum(!is.na(article1tuva$tuva3sbp))
tuva3_n

psych::describe(article1tuva$tuva3sbp)


#TUVA4
tuva4_n <- sum(!is.na(article1tuva$tuva4sbp))
tuva4_n

psych::describe(article1tuva$tuva4sbp)


#TUVA5
tuva5_n <- sum(!is.na(article1tuva$tuva5sbp))
tuva5_n

psych::describe(article1tuva$tuva5sbp)

 


#UTUVA1
utuva1_n <- sum(!is.na(article1utuva$utuva1sbp))
utuva1_n

psych::describe(article1utuva$utuva1sbp)



#UTUVA2
utuva2_n <- sum(!is.na(article1utuva$utuva2sbp))
utuva2_n

psych::describe(article1utuva$utuva2sbp)



 
#UTUVA3
utuva3_n <- sum(!is.na(article1utuva$utuva3sbp))
utuva3_n

psych::describe(article1utuva$utuva3sbp)

```



# Count mean years of the visits in different time points of TUVA and UTUVA
<br><br>

<br><br><br>
```{r mean_year_of_TUVA1, echo=FALSE}


tuva1_year <- year(mean(article1tuva$tuva1_date, na.rm = TRUE))
tuva2_year <- year(mean(article1tuva$tuva2_date, na.rm = TRUE))
tuva3_year <- year(mean(article1tuva$tuva3_date, na.rm = TRUE))
tuva4_year <- year(mean(article1tuva$tuva4_date, na.rm = TRUE))
tuva5_year <- year(mean(article1tuva$tuva5_date, na.rm = TRUE))

utuva1_year <- year(mean(article1tuva$tuva1_date, na.rm = TRUE))
utuva2_year <- year(mean(article1tuva$tuva2_date, na.rm = TRUE))
utuva3_year <- year(mean(article1tuva$tuva3_date, na.rm = TRUE))


```
<br><br><br>

 
 


# The mean SBPs of TUVA and UTUVA time points with the custom function mean_sd()
<br><br>

Pay notice that the `na.rm = TRUE` has to be present. Otherwise the `mean()` function gives only `NA`.

 

```{r SBP_TUVA1}

tuva1_sbp <- mean_sd(article1tuva$tuva1sbp, na.rm = TRUE)
tuva1_sbp

tuva2_sbp <- mean_sd(article1tuva$tuva2sbp, na.rm = TRUE)
tuva2_sbp


tuva3_sbp <- mean_sd(article1tuva$tuva3sbp, na.rm = TRUE)
tuva3_sbp


tuva4_sbp <- mean_sd(article1tuva$tuva4sbp, na.rm = TRUE)
tuva4_sbp
  

tuva5_sbp <- mean_sd(article1tuva$tuva5sbp, na.rm = TRUE)
tuva5_sbp

 
utuva1_sbp <- mean_sd(article1utuva$utuva1sbp, na.rm = TRUE)
utuva1_sbp
 

utuva2_sbp <- mean_sd(article1utuva$utuva2sbp, na.rm = TRUE)
utuva2_sbp


utuva3_sbp <- mean_sd(article1utuva$utuva3sbp, na.rm = TRUE)
utuva3_sbp



```


 
 

# The mean ages in TUVA and UTUVA time points
<br><br>

Pay notice that the `na.rm = TRUE` has to be present. Otherwise the `mean()` function gives only `NA`.

 

```{r age_TUVA1}


tuva1_age <-  article1tuva %>% drop_na(tuva1sbp) %>% select(tuva1age) %>% unlist() %>% mean_sd(., na.rm = T)
tuva1_age
article1tuva %>% drop_na(tuva1sbp) %>% nrow()



tuva2_age <-  article1tuva %>% drop_na(tuva2sbp) %>% select(tuva2age) %>% unlist() %>% mean_sd(., na.rm = T)
tuva2_age
article1tuva %>% drop_na(tuva2sbp) %>% nrow()


tuva3_age <-  article1tuva %>% drop_na(tuva3sbp) %>% select(tuva3age) %>% unlist() %>% mean_sd(., na.rm = T)
tuva3_age
article1tuva %>% drop_na(tuva3sbp) %>% nrow()

article1tuva %>% drop_na(tuva3sbp) %>% select(tuva3age) %>% psych::describe(.)


tuva4_age <-  article1tuva %>% drop_na(tuva4sbp) %>% select(tuva4age) %>% unlist() %>% mean_sd(., na.rm = T)
tuva4_age
article1tuva %>% drop_na(tuva4sbp) %>% nrow()



tuva5_age <-  article1tuva %>% drop_na(tuva5sbp) %>% select(tuva5age) %>% unlist() %>% mean_sd(., na.rm = T)
tuva5_age
article1tuva %>% drop_na(tuva5sbp) %>% nrow()



utuva1_age <-  article1utuva %>% drop_na(utuva1sbp) %>% select(utuva1age) %>% unlist() %>% mean_sd(., na.rm = T)
utuva1_age
article1utuva %>% drop_na(utuva1sbp) %>% nrow()



utuva2_age <-  article1utuva %>% drop_na(utuva2sbp) %>% select(utuva2age) %>% unlist() %>% mean_sd(., na.rm = T)
utuva2_age
article1utuva %>% drop_na(utuva2sbp) %>% nrow()



utuva3_age <-  article1utuva %>% drop_na(utuva3sbp) %>% select(utuva3age) %>% unlist() %>% mean_sd(., na.rm = T)
utuva3_age
article1utuva %>% drop_na(utuva3sbp) %>% nrow()


```

 

 

# Within cohort comparisons with the paired t-test
<br><br>

Now we run the t-tests.

**It is utterly important to pay attention that these has to be paired t-tests, since we are comparing the SBPs from same individuals measured in two different time points.**


Paired t-test TUVA1 vs TUVA2:

<br><br><br>
```{r t-test_SBPs_TUVA1_TUVA2}

# Run model
tuva1_vs_tuva2 <-t.test(article1tuva$tuva1sbp, article1tuva$tuva2sbp, paired=TRUE)
tuva1_vs_tuva2

# Extract p-value
tuva1_vs_tuva2_p <- tuva1_vs_tuva2$p.value
tuva1_vs_tuva2_p

# Stylize
tuva1_vs_tuva2_p <- stylize_p(tuva1_vs_tuva2_p)

```
<br><br><br>

 
 
 Paired t-test TUVA1 vs TUVA3:

 
<br><br><br>
```{r t-test_SBPs_TUVA1_TUVA3}

# Run model
tuva1_vs_tuva3 <-t.test(article1tuva$tuva1sbp, article1tuva$tuva3sbp, paired=TRUE)
tuva1_vs_tuva3

# Extract p-value
tuva1_vs_tuva3_p <- tuva1_vs_tuva3$p.value
tuva1_vs_tuva3_p

# Stylize
tuva1_vs_tuva3_p <- stylize_p(tuva1_vs_tuva3_p)

```
<br><br><br>

Paired t-test TUVA1 vs TUVA4:


```{r t-test_SBPs_TUVA1_TUVA4}

 
# Run model
tuva1_vs_tuva4 <-t.test(article1tuva$tuva1sbp,  article1tuva$tuva4sbp, paired=TRUE)
tuva1_vs_tuva4

# Extract p-value
tuva1_vs_tuva4_p <- tuva1_vs_tuva4$p.value
tuva1_vs_tuva4_p

# Stylize
tuva1_vs_tuva4_p <- stylize_p(tuva1_vs_tuva4_p)

```
<br><br><br>

Paired t-test TUVA1 vs TUVA5:

```{r t-test_SBPs_TUVA1_TUVA5}

 
# Run model
tuva1_vs_tuva5 <-t.test(article1tuva$tuva1sbp, article1tuva$tuva5sbp, paired=TRUE)
tuva1_vs_tuva5

# Extract p-value
tuva1_vs_tuva5_p <- tuva1_vs_tuva5$p.value
tuva1_vs_tuva5_p

# Stylize
tuva1_vs_tuva5_p <- stylize_p(tuva1_vs_tuva5_p)

```

Paired t-test UTUVA1 vs UTUVA2:



```{r t-test_SBPs_UTUVA1_UTUVA2}
 
 
# Run model
utuva1_vs_utuva2 <-t.test(article1utuva$utuva1sbp, article1utuva$utuva2sbp, paired=TRUE)

utuva1_vs_utuva2

# Extract p-value
utuva1_vs_utuva2_p <- utuva1_vs_utuva2$p.value
utuva1_vs_utuva2_p

# Stylize
utuva1_vs_utuva2_p <- stylize_p(utuva1_vs_utuva2_p)


```


Paired t-test UTUVA1 vs UTUVA3:


```{r t-test_SBPs_UTUVA1_UTUVA3}
 
  
# Run model
utuva1_vs_utuva3 <-t.test(article1utuva$utuva1sbp, article1utuva$utuva3sbp, paired=TRUE)

utuva1_vs_utuva3

# Extract p-value
utuva1_vs_utuva3_p <- utuva1_vs_utuva3$p.value
utuva1_vs_utuva3_p
 
# Stylize
utuva1_vs_utuva3_p <- stylize_p(utuva1_vs_utuva3_p)


```


# Comparison of SBPs between cohorts with un-paired t-tests
<br><br>

Let us start by counting:

1. the mean SBPs for TUVA1 and UTUVA1
2. their subtraction

<br><br><br>
```{r subtraction_SBPs_TUVA1_UTUVA1}
 
tuva1_utuva1_subtraction <- mean(article1tuva$tuva1sbp, na.rm = TRUE) - mean(article1utuva$utuva1sbp, na.rm = TRUE)

```
<br><br><br>
 
Thereafter, let us run the un-paired t-test.

(paired=FALSE)
 

<br><br><br>
```{r t-test_SBPs_TUVA1_UTUVA1}
 
# Run model
tuva1_vs_utuva1 <-t.test(article1tuva$tuva1sbp, article1utuva$utuva1sbp, paired=FALSE)

tuva1_vs_utuva1

# Extract p-value
tuva1_vs_utuva1_p <- tuva1_vs_utuva1$p.value
tuva1_vs_utuva1_p

# Stylize
tuva1_vs_utuva1_p <- stylize_p(tuva1_vs_utuva1_p)
  
```
<br><br><br>


Subtraction of means of TUVA2 and UTUVA3

<br><br><br>
```{r subtraction_SBPs_TUVA2_UTUVA3}
 
tuva2_utuva3_subtraction <- mean(article1tuva$tuva2sbp, na.rm = TRUE) - mean(article1utuva$utuva3sbp, na.rm = TRUE)

```
<br><br><br>

t-test UTUVA3 vs TUVA2:

```{r t-test_SBPs_TUVA2_UTUVA3}
 
# Run model
tuva2_vs_utuva3 <-t.test(article1tuva$tuva2sbp, article1utuva$utuva3sbp, paired=FALSE)

tuva2_vs_utuva3

# Extract p-value
tuva2_vs_utuva3_p <- tuva2_vs_utuva3$p.value
tuva2_vs_utuva3_p

# Stylize
tuva2_vs_utuva3_p <- stylize_p(tuva2_vs_utuva3_p)

```





# Compose Table 2 of the results


<br><br><br>
```{r table2}
 
##################################################################
##                             TUVA                             ##
##################################################################


table2_part1 <- 
tibble::tribble(
  ~Cohort,    ~I, ~II,  ~III,   ~IV,    ~V,   ~VI,
  "Year",  tuva1_year, 0, tuva2_year, tuva3_year, tuva4_year, tuva5_year,
    "n", tuva1_n, 0, tuva2_n, tuva3_n, tuva4_n, tuva5_n,
  )# %>% as.data.frame()

table2_part1 <- table2_part1 %>% 
  mutate(across(where(is.numeric), ~ round_half_up(., digits = 0))) %>% 
  mutate(across(where(is.numeric), ~ as.character(.)))


table2_part2a <- 
tibble::tribble(
  ~Cohort,    ~I, ~II,  ~III,   ~IV,    ~V,   ~VI,
    "Mean age (years)", tuva1_age, NA,  tuva2_age, tuva3_age, tuva4_age, tuva5_age
  ) #%>% as.data.frame()

# table2_part2a <- table2_part2a %>%
#   mutate(across(where(is.numeric), ~ round_half_up(., digits = 1))) %>%
#   mutate(across(where(is.numeric), ~ as.character(.)))

table2_part2b <- 
tibble::tribble(
  ~Cohort,    ~I, ~II,  ~III,   ~IV,    ~V,   ~VI,
   "SBP (mmHg)", tuva1_sbp, 0, tuva2_sbp, tuva3_sbp, tuva4_sbp, tuva5_sbp
  ) #%>% as.data.frame()




table2_part3 <- 
tibble::tribble(
  ~Cohort,    ~I, ~II,  ~III,   ~IV,    ~V,   ~VI,
   "p-value", "", "", tuva1_vs_tuva2_p, tuva1_vs_tuva3_p, tuva1_vs_tuva4_p, tuva1_vs_tuva5_p
  )


#################################################################
##                            UTUVA                            ##
#################################################################

table2_part4 <- 
tibble::tribble(
  ~Cohort,    ~I, ~II,  ~III,   ~IV,    ~V,   ~VI,
  "Year",  utuva1_year, utuva2_year, utuva3_year, NA, NA, NA,
    "n", utuva1_n, utuva2_n, utuva3_n, NA, NA, NA
  )

table2_part4 <- table2_part4 %>% 
  mutate(across(where(is.numeric), ~ round_half_up(., digits = 0))) %>% 
  mutate(across(where(is.numeric), ~ as.character(.)))




table2_part5a <- 
tibble::tribble(
  ~Cohort,    ~I, ~II,  ~III,   ~IV,    ~V,   ~VI,
  
  "Mean age (years)", utuva1_age, utuva2_age, utuva3_age,  NA, NA, NA

  )  

# table2_part5a <- table2_part5a %>% 
#   mutate(across(where(is.numeric), ~ round_half_up(., digits = 1))) %>% 
#   mutate(across(where(is.numeric), ~ as.character(.)))



table2_part5b <- 
tibble::tribble(
  ~Cohort,    ~I, ~II,  ~III,   ~IV,    ~V,   ~VI,

     "SBP (mmHg)", utuva1_sbp, utuva2_sbp, utuva3_sbp, NA, NA, NA
  )  




table2_part6 <- 
tibble::tribble(
  ~Cohort,    ~I, ~II,  ~III,   ~IV,    ~V,   ~VI,
   "p-value", "", utuva1_vs_utuva2_p, utuva1_vs_utuva3_p, "", "", ""
  )

 
table2_part7 <- 
tibble::tribble(
  ~Cohort,    ~I, ~II,  ~III,   ~IV,    ~V,   ~VI,
   "Difference in SBP between cohorts (mmHg)", tuva1_utuva1_subtraction, "", tuva2_utuva3_subtraction, "", "", ""
  )

 
table2_part7 <- table2_part7 %>% 
  mutate(across(where(is.numeric), ~ round_half_up(., digits = 1))) %>% 
  mutate(across(where(is.numeric), ~ as.character(.)))

 

table2_part8 <- 
tibble::tribble(
  ~Cohort,    ~I, ~II,  ~III,   ~IV,    ~V,   ~VI,
   "p-value", tuva1_vs_utuva1_p, "", tuva2_vs_utuva3_p, "", "", ""
  )

table2 <- 
rbind(table2_part1,
table2_part2a,
table2_part2b,
table2_part3,
table2_part4,
table2_part5a,
table2_part5b,
table2_part6,
table2_part7,
table2_part8)


table2 <- table2 %>% add_row(Cohort="TUVA",
                             I="",
                             II="",
                             III="",
                             IV="",
                             V="",
                             VI="",
                           .before=1)

 table2 <- table2 %>% add_row(Cohort="UTUVA",
                              I="",
                              II="",
                              III="",
                              IV="",
                              V="",
                              VI="",
                            .before=7)




table2 %>% kable() %>% kable_styling()


```
<br><br><br>
 





# Export to excel

```{r exportexcel}
  

# Export to Excel file
write_xlsx(table2, "article1_sensitivity_analyses_table2_sbp.xlsx")



```



