---
title: "The Natural History of Hypertension in Older Adults: A Study of Two Finnish Generational Cohorts Born 20 Years Apart - supplemental figure, systolic BP analyses"
output: html_document
classoption: landscape
date: "`r Sys.Date()`"
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages

# First pacman
library(pacman)

# Load the rest
p_load(ggplot2, ggthemes, dplyr, knitr, kableExtra, here, tidyverse,
       magick, showtext, lubridate)

# Tell the "here" package where you are right now - i.e. tell the location of the file you are running right now

here::i_am("article1/README.R")

# After that, you can check where the project root is by running "here()"

here()

#  Load the datasets using {here} package
article1tuva_file <- here("data", "a1rev_tuva.rds")
 
article1tuva <- readRDS(article1tuva_file)
 
# Clean up variables
rm(article1tuva_file)


# Format the kable styling
bs_style <- c("striped", "bordered")

options(kable_styling_bootstrap_options = bs_style)



```
<br><br><br><br>
 



# Exclusion of participants in the baseline not surviving to the 95-year-old's timepoint
<br><br>


TUVA1:


```{r exclusion_tuva1, echo=FALSE}

# TUVA1
article1tuva <- article1tuva %>% drop_na(tuva5sbp, tuva5dbp)

# Display the n of this supplemental figure
nrow(article1tuva)

```

 

# Introduce a custom function to calculate 95% confidence intervals

There is no ready function in R to calculate 95% CIs. However, such a custom function can be easily made. We only need a function for the upper bound (lower can be calculated later on the spot).
 
<br>
```{r custom_function}

ci_95_upper <- function(x) {
  ci <- t.test(x, na.action = na.omit)$conf.int
  upper_bound <- ci[2]
  return(upper_bound)
}
 

```
<br><br><br>



# Introduce a custom function to a mean year from a date variable

There is no ready function in R to calculate 95% CIs. However, such a custom function can be easily made. We only need a function for the upper bound (lower can be calculated later on the spot).
 
<br>
```{r mean_year}

# Define the function
mean_year <- function(date_variable) {
  # Calculate the mean of the date variable, ignoring NA values
  mean_date_numeric <- mean(as.numeric(date_variable), na.rm = TRUE)
  
  # Convert the numeric mean date to a Date object
  mean_date <- as.Date(mean_date_numeric, origin = "1970-01-01")
  
  # Extract and return the year from the mean date
  return(year(mean_date))
} 

```
<br><br><br>

 

 
# Calculate means and 95% confidence intervals for plotting

Note that only one of the 95% CI bounds is needed for the plot (the other can be calculated on the spot)



## For the supplemental figure, only the SBP readings of those reaching 95 years in TUVA are needed

<br>
```{r statistics_tuva_part1}



# Calculate mean and SD of tuva1
SBP70_tuva_mean <- mean(article1tuva$tuva1sbp, na.rm = T)
SBP70_tuva_mean

SBP70_tuva_upper <- ci_95_upper(article1tuva$tuva1sbp)
SBP70_tuva_upper



# Mean and SD of tuva2
SBP80_tuva_mean <- mean(article1tuva$tuva2sbp, na.rm = T)
SBP80_tuva_mean

SBP80_tuva_upper <- ci_95_upper(article1tuva$tuva2sbp)
SBP80_tuva_upper



# Mean and SD of tuva3
SBP85_tuva_mean <- mean(article1tuva$tuva3sbp, na.rm = T)
SBP85_tuva_mean

SBP85_tuva_upper <- ci_95_upper(article1tuva$tuva3sbp)
SBP85_tuva_upper



# Mean of tuva4
SBP90_tuva_mean <- mean(article1tuva$tuva4sbp, na.rm = T)
SBP90_tuva_mean

SBP90_tuva_upper <- ci_95_upper(article1tuva$tuva4sbp)
SBP90_tuva_upper

# Mean and SD of tuva5
SBP95_tuva_mean <- mean(article1tuva$tuva5sbp, na.rm = T)
SBP95_tuva_mean

SBP95_tuva_upper <- ci_95_upper(article1tuva$tuva5sbp)
SBP95_tuva_upper



```





## Calculate the mean ages in each timepoint

<br>
```{r mean_ages}


tuva1_age <-  article1tuva %>% drop_na(tuva1sbp) %>% select(tuva1age) %>% unlist() %>% mean(., na.rm = T)
tuva1_age
article1tuva %>% drop_na(tuva1sbp) %>% nrow()



tuva2_age <-  article1tuva %>% drop_na(tuva2sbp) %>% select(tuva2age) %>% unlist() %>% mean(., na.rm = T)
tuva2_age
article1tuva %>% drop_na(tuva2sbp) %>% nrow()



tuva3_age <-  article1tuva %>% drop_na(tuva3sbp) %>% select(tuva3age) %>% unlist() %>% mean(., na.rm = T)
tuva3_age
article1tuva %>% drop_na(tuva3sbp) %>% nrow()



tuva4_age <-  article1tuva %>% drop_na(tuva4sbp) %>% select(tuva4age) %>% unlist() %>% mean(., na.rm = T)
tuva4_age
article1tuva %>% drop_na(tuva4sbp) %>% nrow()



tuva5_age <-  article1tuva %>% drop_na(tuva5sbp) %>% select(tuva5age) %>% unlist() %>% mean(., na.rm = T)
tuva5_age
article1tuva %>% drop_na(tuva5sbp) %>% nrow()




```





# Store the statistics that were created above to a custom data frame (to a tribble)

<br>
```{r tribble}



data_for_plot <- 

tibble::tribble(
  ~Age, ~Cohort,          ~SBP_mean,      ~upperCI,
   tuva1_age,  "1920-born TUVA cohort",  SBP70_tuva_mean,  SBP70_tuva_upper,
   tuva2_age,  "1920-born TUVA cohort",  SBP80_tuva_mean,  SBP80_tuva_upper,
   tuva3_age,  "1920-born TUVA cohort",  SBP85_tuva_mean,  SBP85_tuva_upper,
   tuva4_age,  "1920-born TUVA cohort",  SBP90_tuva_mean,  SBP90_tuva_upper,
   tuva5_age,  "1920-born TUVA cohort",  SBP95_tuva_mean,  SBP95_tuva_upper

  )

data_for_plot <- data_for_plot %>% mutate(lowerCI = SBP_mean-upperCI+SBP_mean)

uplabel1 <- data_for_plot$upperCI[1] + 2
uplabel2 <- data_for_plot$upperCI[2] + 2
uplabel3 <- data_for_plot$upperCI[3] + 2
uplabel4 <- data_for_plot$upperCI[4] + 2
uplabel5 <- data_for_plot$upperCI[5] + 2



```
<br><br><br>


# Calculate mean years


```{r}

# TUVA
tuva1year <- mean_year(article1tuva$tuva1_date)
tuva2year <- mean_year(article1tuva$tuva2_date)
tuva3year <- mean_year(article1tuva$tuva3_date)
tuva4year <- mean_year(article1tuva$tuva4_date)
tuva5year <- mean_year(article1tuva$tuva5_date)



```

# Finally, create the plot
 
<br>
```{r fig1, echo=FALSE}


# Load specific font from Google Fonts
font_add_google("Rosario", family = "rosario")

# Invoke showtext
showtext_auto()

# Define your custom colors

custom_colors <- c("1920-born TUVA cohort" = "#5C7DB2", "1940-born UTUVA cohort" = "#E29A3F")
shape_mapping <- c("1920-born TUVA cohort" = 19, "1940-born UTUVA cohort" = 15)  


fig1 <-
ggplot(data_for_plot, aes(x = Age, y = SBP_mean, color = Cohort, shape = Cohort)) +
  
  geom_line(linewidth = 0.8) +

  geom_point(size = 2.5) +
  
  scale_shape_manual(values = shape_mapping) + # Apply the shape mapping

  geom_errorbar(aes(ymin=SBP_mean-upperCI+SBP_mean,
                    ymax=upperCI),
                width=0.4,
                size=0.8,
                ) +  
 
  scale_color_manual(values = custom_colors) +  # Manually set fill colors

   
  labs(x = "Age", y = "SBP mean") +

  theme_classic(base_size = 15, base_family = "rosario") +  
    

  scale_x_continuous(breaks = seq(70, 95, 5),
                     name = "Age (years)") +
  scale_y_continuous(breaks = seq(140, 190, 10), limits = c(140, 190),
                     name = "Systolic blood pressure (mmHg)",
                     expand = c(0, 0)) +

  
   theme(axis.title.x = element_text(face = "bold",
    margin = margin(t = 15, unit = "pt")),  # Adjust margin for x-axis title
    axis.title.y = element_text(face = "bold",
      margin = margin(r = 15, unit = "pt"))) +  # Adjust margin for y-axis title

  

# The year labels of TUVA:
annotate("text", x=tuva1_age, y=uplabel1, label=tuva1year, color="black", size=4,
               family = "rosario") +
annotate("text", x=tuva2_age, y=uplabel2, label=tuva2year, color="black", size=4,
               family = "rosario") +
annotate("text", x=tuva3_age, y=uplabel3, label=tuva3year, color="black", size=4,
               family = "rosario") +
annotate("text", x=tuva4_age, y=uplabel4, label=tuva4year, color="black", size=4,
               family = "rosario") +
annotate("text", x=tuva5_age, y=uplabel5, label=tuva5year, color="black", size=4,
               family = "rosario") +


# The n label:
annotate("text", x=70, y=189, label="Only participants reaching", color="black", size=4, hjust = 0,
               family = "rosario") +
annotate("text", x=70, y=186, label="95-year-olds' evaluation", color="black", size=4, hjust = 0,
               family = "rosario") +
annotate("text", x=70, y=183, label="in TUVA study (n=55)", color="black", size=4, hjust = 0,
               family = "rosario") +

#geom_rect(aes(xmin = 84, xmax = 94, ymin = 136, ymax = 148), 
#            color = "black", fill = NA, linetype = "solid", size = 0.3)  +
  

theme(legend.position = "none")




fig1

```
<br><br><br>



# Export the fig. 1 to PDF and PNG

We will conclude this R Markdown file by exporting the Fig. 1 to a PDF and PNG file.



```{r export, echo=FALSE}

# Save as PDF with dpi specified
ggsave("Fig_S1_SBP_dpi600_revision.pdf", fig1, width = 7, height =5, dpi = 600)


# Load that pdf file with the magick package
pdf_image <- magick::image_read_pdf("Fig_S1_SBP_dpi600_revision.pdf", density = 600)

# Save it as PNG
image_write(pdf_image,
            path = "Fig_S1_SBP_dpi600_revision.png",
            format = "png",
            density = 600)


```

 
